---
title: "NP_1 seurat practice"
author: "Lifei Jiang"
date: "2/16/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup the Seurat Object

##Load the data and initialize seurat
```{r}
library(dplyr)
library(Seurat)

#load
NP1.data <- Read10X(data.dir = "~/data/NP_1/outs/filtered_feature_bc_matrix/")
nrow(NP1.data) #31053
#initialize seurat
NP1 <- CreateSeuratObject(counts = NP1.data, project = "NP1project", assay = "RNA", min.cells = 3, min.features = 200)

NP1

```
## Explore the data: find that each row is a gene, and each colomn is a cell, represented by a barcode 
```{r}
nrow(NP1.data)
ncol(NP1.data)
NP1.data[c("Xkr4", "Gm1992", "Gm37381"), 1:3] #First run NP1.data to see the gene name
```



# Filter cells:QC and selecting cells for further analysis
```{r}

#add a mt gene percentage colomn, calulate the percentage of MT genes in each cell
NP1[["percent.mt"]] <- PercentageFeatureSet(NP1, pattern = "^mt-")

VlnPlot(NP1, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

#mt<7%
#nFeature>500, nFeature< 6000

plot1 <- FeatureScatter(NP1, feature1 = "nFeature_RNA", feature2 = "percent.mt")
plot1
plot2 <- FeatureScatter(NP1, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot2
CombinePlots(plots = list(plot1, plot2))

NP1 <- subset(NP1, subset = nFeature_RNA > 500 & nFeature_RNA < 6000 & percent.mt < 7)

```

# Normalizing the data
```{r}

# normalizes the feature expression measurements for each cell by the total expression
# LogNormalize: Feature counts for each cell are divided by the total counts for that cell and multiplied by the scale.factor. This is then natural-log transformed using log1p.
NP1 <- NormalizeData(NP1, normalization.method = "LogNormalize", scale.factor = 10000)
```

# Identification of highly variable features (feature selection)
```{r}
NP1 <- FindVariableFeatures(NP1, selection.method = "vst", nfeatures = 2000)

## Use “attributes(NP1)” to see the identified VariableFeatures

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(NP1), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(NP1)
plot1
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot2
CombinePlots(plots = list(plot1, plot2))
```

# Scaling the data for PCA

- Shifts the expression of each gene, so that the mean expression across cells is 0
- Scales the expression of each gene, so that the variance across cells is 1

```{r}
all.genes <- rownames(NP1)
NP1 <- ScaleData(NP1, features = all.genes)
```

# Run PCA

```{r}
NP1 <- RunPCA(NP1, features = VariableFeatures(object = NP1))

ElbowPlot(NP1)

## choose PC1-PC10
```


# Cluster the cells 

```{r}
NP1 <- FindNeighbors(NP1, dims = 1:9)
NP1 <- FindClusters(NP1, resolution = 0.03)
##?? how much resolution should I use? We find that setting this parameter between 0.4-1.2 typically returns good results for single-cell datasets of around 3K cells. Optimal resolution often increases for larger datasets.

## First need to use low resolution to identify the big clusters, then classify subsets within each group. 
head(Idents(NP1), 5)

#Number of communities: 6

```


# Run non-linear dimensional reduction (UMAP)
```{r}
NP1 <- RunUMAP(NP1, dims = 1:9)

DimPlot(NP1, reduction = "umap")
```


# Run non-linear dimensional reduction (tSNE)

```{r}
NP1 <- RunTSNE(NP1, dims = 1:9)

DimPlot(NP1, reduction = "tsne")

# Should mainly have three populations: myoepithelial，progenitor，luminal; first identify the marker for each big cluster --> then classify within each cluster

```


# Save the object, no need to return later (the end of the file address should be the name of this file)
```{r}
saveRDS(NP1, file = "~/data/NP1_tutorial.rds")
```




```{r}
NP1 <- readRDS("/home/jianglf/data/NP1_tutorial.rds")
```

# Find markers for each cluster
```{r}
# Find markers for each cluster
cluster0.markers <- FindMarkers(NP1, ident.1 = 0, min.pct = 0.25, logfc.threshold = 0.4)

cluster0.markers

write.csv(cluster0.markers, file = "~/Lifei Jiang R code/cluster0.markers.csv")


# Filename also contain i, need to combine characters and numbers, first define filename

for(i in 0:3){
 filenames=paste("~/refined_GeneMarkers/cluster", i, ".markers.csv", sep="") 
 write.csv(FindMarkers(NP1, ident.1=i, min.pct=0.25,logfc.threshold = 1, only.pos = TRUE) ,filenames)
}

# explore the paste function 
paste("~/Lifei Jiang R code/cluster", 1, ".markers.csv", sep="")
paste("A", 1:6, sep="")

# Find all markers for each cluster compared to all the other clusters. only the positive ones
# Test whether the marker is good, test.use="roc", close to 1 is perfect


# library(dplyr)
NP1.markers <- FindAllMarkers(NP1, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
NP1.markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_logFC)
```

# Plot the logfc and p value distribution for cluster markers
```{r}
# 4 clusters (3 main cluster, 1 rare clusters)
 
mlmarkers <- read.csv (file = "~/refined_GeneMarkers/cluster0.markers.csv")
mlmarkers <- mlmarkers[mlmarkers$p_val_adj<0.05, ]
str(mlmarkers)#1059
mlmarkers_Pos <- as.character(mlmarkers[mlmarkers$avg_logFC>0, ]$X)
basalmarkers <- read.csv (file = "~/refined_GeneMarkers/cluster1.markers.csv")
basalmarkers <- basalmarkers[basalmarkers$p_val_adj<0.05, ]
str(basalmarkers)#1139
basalmarkers_Pos <- as.character(basalmarkers[basalmarkers$avg_logFC>0, ]$X)
lpmarkers <- read.csv (file = "~/refined_GeneMarkers/cluster2.markers.csv")
lpmarkers <- lpmarkers[lpmarkers$p_val_adj<0.05, ]
str(lpmarkers) #671
lpmarkers_Pos <-as.character(lpmarkers[lpmarkers$avg_logFC>0, ]$X)

raremarkers <- read.csv (file = "~/refined_GeneMarkers/cluster3.markers.csv")
raremarkers <- raremarkers[raremarkers$p_val_adj<0.05, ]
str(raremarkers) #671
raremarkers_Pos <-as.character(raremarkers[raremarkers$avg_logFC>0, ]$X)


plot(mlmarkers$p_val_adj, mlmarkers$avg_logFC, main="mlmarkers") #p=0.05, logfc=0.2
plot(basalmarkers$p_val_adj, basalmarkers$avg_logFC, main="basalmarkers") #p=0.05, logfc=0.2
plot(lpmarkers$p_val_adj, lpmarkers$avg_logFC, main="lpmarkers") #p=0.05, logfc=0.2

```

#Using heatmap to check the specificity of the data
```{r}
#  Create a data object for cluster markers 
MammaryMarkers <- list("ML"=mlmarkers_Pos,"Basal"=basalmarkers_Pos,"LP"= lpmarkers_Pos,"Rare"=raremarkers_Pos) # using logfc threshold = 1

saveRDS(MammaryMarkers, file="~/data/MammaryMarkers.rds")


features.4clusters <- c(mlmarkers_Pos, basalmarkers_Pos, lpmarkers_Pos, raremarkers_Pos)
DoHeatmap(NP1, features = features.4clusters) + NoLegend()
```

#Annotation for the big clusters
```{r}

new.clusterbig.ids <- c("ML", "Basal","LP","Rare1")
names(new.clusterbig.ids) <- levels(NP1)
NP1<-RenameIdents(NP1, new.clusterbig.ids)


DimPlot(NP1, reduction = "umap", label = TRUE, pt.size = 0.5) + NoLegend()
```

# First save the current identity

```{r}

NP1[["ClusterNames_0.1"]] <- Idents(object = NP1)

```
# Save the big clusters 
```{r}
saveRDS(NP1, file = "~/data/NP1_tutorial.2.rds")
```

## Subsetting the data for basal 
```{r}
#### ??? 


NP1.basal <- subset(NP1,idents="Basal")

DietSeurat(NP1.basal, counts = TRUE, data = TRUE, scale.data = FALSE)

```

# Clusters and markers within the basal group 
```{r}
# Quality control 
# Filter cells:QC and selecting cells for further analysis


#add a mt gene percentage colomn, calulate the percentage of MT genes in each cell
NP1.basal[["percent.mt"]] <- PercentageFeatureSet(NP1.basal, pattern = "^mt-")

VlnPlot(NP1.basal, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

#mt<7%
#nFeature>500, nFeature< 6000

plot1 <- FeatureScatter(NP1.basal, feature1 = "nFeature_RNA", feature2 = "percent.mt")
plot1
plot2 <- FeatureScatter(NP1.basal, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot2
CombinePlots(plots = list(plot1, plot2))

NP1.basal <- subset(NP1.basal, subset = nFeature_RNA > 500 & nFeature_RNA < 6000 & percent.mt < 7)

# Normalization 
NP1.basal <- NormalizeData(NP1.basal, normalization.method = "LogNormalize", scale.factor = 1000000)

# Variable feature identification

NP1.basal <- FindVariableFeatures(NP1.basal, selection.method = "vst", nfeatures = 2000)

## Use “attributes(NP1.basal)” to see the identified VariableFeatures

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(NP1.basal), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(NP1.basal)
plot1
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot2
CombinePlots(plots = list(plot1, plot2))


# Scale the data 
all.genes <- rownames(NP1.basal)
NP1.basal <- ScaleData(NP1.basal, features = all.genes)


# Run PCA

NP1.basal <- RunPCA(NP1.basal, features = VariableFeatures(object = NP1.basal))

ElbowPlot(NP1.basal)

## choose PC1-PC10



NP1.basal <- FindNeighbors(NP1.basal, dims = 1:9)
NP1.basal <- FindClusters(NP1.basal, resolution = 0.2)
##?? how much resolution should I use? We find that setting this parameter between 0.4-1.2 typically returns good results for single-cell datasets of around 3K cells. Optimal resolution often increases for larger datasets.

## First need to use low resolution to identify the big clusters, then classify subsets within each group. 
head(Idents(NP1.basal), 5)


NP1.basal <- RunUMAP(NP1.basal, dims = 1:9)

DimPlot(NP1.basal, reduction = "umap")


## Identify cluster markers 
Basal.subset <- FindAllMarkers(NP1.basal,only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)

Basal.subset <- Basal.subset[Basal.subset$p_val_adj<0.05, ]

Basal.0 <- Basal.subset[Basal.subset$cluster=="0",]$gene
Basal.1 <- Basal.subset[Basal.subset$cluster=="1",]$gene
Basal.2 <- Basal.subset[Basal.subset$cluster=="2",]$gene
Basal.3 <- Basal.subset[Basal.subset$cluster=="3",]$gene


# Quality of identified markers
Basal.subset <- FindAllMarkers(NP1.basal,only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
DoHeatmap(NP1.basal, features = Basal.subset$gene) + NoLegend()


saveRDS(NP1.basal, file="~/data/NP1.basal.rds")
```

## Find nonoverlapping markers
```{r}
# Load the library
library(limma)


# What are the possible letters in the universe?
universe <- sort(unique(c(Basal.0, Basal.1, Basal.2, Basal.3)))

# Generate a matrix, with the sets in columns and possible letters on rows
Counts <- matrix(0, nrow=length(universe), ncol=4)
# Populate the said matrix
for (i in 1:length(universe)) {
   Counts[i,1] <- universe[i] %in% Basal.0
   Counts[i,2] <- universe[i] %in% Basal.1
   Counts[i,3] <- universe[i] %in% Basal.2
   Counts[i,4] <- universe[i] %in% Basal.3
}

# Name the columns with the sample names
colnames(Counts) <- c("Basal.0", "Basal.1", "Basal.2", "Basal.3")

# Specify the colors for the sets
cols<-c("Red", "Green", "Blue","Grey")
vennDiagram(vennCounts(Counts), circle.col=cols)




## Selecting the nonoverlapping genes from each group 

Basal.0 <- Basal.0[which(Basal.0 %in% c(Basal.1, Basal.2, Basal.3) == FALSE)]

write.csv(Basal.0, file = "~/Lifei Jiang R code/Basal0.csv")

Basal.1 <- Basal.1[which(Basal.1 %in% c(Basal.0, Basal.2, Basal.3) == FALSE)]
write.csv(Basal.1, file = "~/Lifei Jiang R code/Basal1.csv")

Basal.2 <- Basal.2[which(Basal.2 %in% c(Basal.1, Basal.0, Basal.3) == FALSE)]
write.csv(Basal.2, file = "~/Lifei Jiang R code/Basal2.csv")

Basal.3 <- Basal.3[which(Basal.3 %in% c(Basal.1, Basal.2, Basal.0) == FALSE)]
write.csv(Basal.3, file = "~/Lifei Jiang R code/Basal3.csv")

Basal.all <- c(Basal.0, Basal.1, Basal.2, Basal.3 )

# Quality of identified markers
DoHeatmap(NP1.basal, features = Basal.all) + NoLegend()
DoHeatmap(NP1.basal, features = Basal.2) + NoLegend()

```

# Improve the classification within the Basal subset 
```{r}
# Since Cluster1 markers are too few, cannot really be seperated from cluster 0, therefore I decided to reduce the resolution to only include 3 clusters 

NP1.basal <- FindNeighbors(NP1.basal, dims = 1:9)
NP1.basal <- FindClusters(NP1.basal, resolution = 0.15)

NP1.basal <- RunUMAP(NP1.basal, dims = 1:9)

DimPlot(NP1.basal, reduction = "umap")

## Identify cluster markers 
Basal.subset <- FindAllMarkers(NP1.basal,only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)

Basal.subset <- Basal.subset[Basal.subset$p_val_adj<0.05, ]

Basal.0 <- Basal.subset[Basal.subset$cluster=="0",]$gene
Basal.1 <- Basal.subset[Basal.subset$cluster=="1",]$gene
Basal.2 <- Basal.subset[Basal.subset$cluster=="2",]$gene



# Quality of identified markers
DoHeatmap(NP1.basal, features = Basal.subset$gene) + NoLegend()

### But I think the heatmap result is no better than 4 clusters 

```


```{r}
readRDS(NP1, file = "~/data/NP1_tutorial.2.rds")
```



```{r}
# Sub clusters within each cluster

# Increase the resolution 
NP1 <- FindClusters(NP1, resolution = 0.8)

# Plot using Umap
NP1 <- RunUMAP(NP1, dims = 1:9)
Idents(NP1)
DimPlot(NP1, reduction = "umap", pt.size = 0.5) + NoLegend()


# Annotation 

new.cluster.ids <- c("Basal1", "ML1", "LP1", "ML2", "ML3", "ML4", "LP2", 
    "ML5", "Basal2", "Basal3","Rare1", "Rare2", "Rare3")
names(new.cluster.ids) <- levels(NP1)
NP1 <- RenameIdents(NP1, new.cluster.ids)
DimPlot(NP1, reduction = "umap", label = TRUE, pt.size = 0.5) + NoLegend()

## Find all markers 
Small.Cluster <- FindAllMarkers(NP1, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
Small.Cluster <- Small.Cluster[Small.Cluster$p_val_adj<0.05, ]

str(Small.Cluster)

Small.Cluster.markers <- Small.Cluster$gene

# Identify the markers for sub-clusters
# 1. basal group

Basal1.markers.3 <- FindMarkers(NP1, ident.1 = "Basal1", ident.2= c("Basal2","Basal3"), min.pct = 0.25,logfc.threshold = 0.2)

write.csv(Basal1.markers.3, file = "~/Lifei Jiang R code/Basal1.markers.3.csv")

Basal2.markers.3 <- FindMarkers(NP1, ident.1 = "Basal2", ident.2= c("Basal1","Basal3"), min.pct = 0.25,logfc.threshold = 0.2)

write.csv(Basal2.markers.3, file = "~/Lifei Jiang R code/Basal2.markers.3.csv")

Basal3.markers.3 <- FindMarkers(NP1, ident.1 = "Basal3", ident.2= c("Basal2","Basal1"), min.pct = 0.25,logfc.threshold = 0.2)

write.csv(Basal3.markers.3, file = "~/Lifei Jiang R code/Basal3.markers.3.csv")

# 2. Mature luminal group: cluster 1,3,4,5,7 

ML1.markers.3 <- FindMarkers(NP1, ident.1 = "ML1", ident.2= c("ML2","ML3", "ML4","ML5"), min.pct = 0.25,logfc.threshold = 0.2)
write.csv(ML1.markers.3, file = "~/Lifei Jiang R code/ML1.markers.3.csv")

ML2.markers.3 <- FindMarkers(NP1, ident.1 = "ML2", ident.2= c("ML1","ML3", "ML4","ML5"), min.pct = 0.25,logfc.threshold = 0.2)
write.csv(ML2.markers.3, file = "~/Lifei Jiang R code/ML2.markers.3.csv")

ML3.markers.3 <- FindMarkers(NP1, ident.1 = "ML3", ident.2= c("ML1","ML2", "ML4","ML5"), min.pct = 0.25,logfc.threshold = 0.2)
write.csv(ML3.markers.3, file = "~/Lifei Jiang R code/ML3.markers.3.csv")

ML4.markers.3 <- FindMarkers(NP1, ident.1 = "ML4", ident.2= c("ML1","ML3", "ML2","ML5"), min.pct = 0.25,logfc.threshold = 0.2)
write.csv(ML4.markers.3, file = "~/Lifei Jiang R code/ML4.markers.3.csv")

ML5.markers.3 <- FindMarkers(NP1, ident.1 = "ML5", ident.2= c("ML1","ML3", "ML4","ML2"), min.pct = 0.25,logfc.threshold = 0.2)
write.csv(ML5.markers.3, file = "~/Lifei Jiang R code/ML5.markers.3.csv")


# 3. Luminal progenitor group: Cluster 2, 6 
LP1.markers.3 <- FindMarkers(NP1, ident.1 = "LP1", ident.2= "LP2", min.pct = 0.25,logfc.threshold = 0.2)
write.csv(LP1.markers.3, file = "~/Lifei Jiang R code/LP1.markers.3.csv")

LP2.markers.3 <- FindMarkers(NP1, ident.1 = "LP2", ident.2= "LP1", min.pct = 0.25,logfc.threshold = 0.2)
write.csv(LP2.markers.3, file = "~/Lifei Jiang R code/LP2.markers.3.csv")

# Save the identity

NP1[["ClusterNames_0.8"]] <- Idents(object = NP1)

# Can use FeaturePlot to visualize the identified markers

# e.g. FeaturePlot(NP1, features = c("Cxcl14", "Slpi", "Ptn", "Wfdc2", "Wfdc18", "Aldh1a3", "Tagln", "Acta2", 
   # "Nupr1"))


```



# Save the small clusters 
```{r}
saveRDS(NP1, file = "~/data/NP1_tutorial.3.rds")
```

## Intermeidate resolutions 
```{r}
# Sub clusters within each cluster

# Increase the resolution 
NP1 <- FindClusters(NP1, resolution = 0.7)

# Plot using Umap
NP1 <- RunUMAP(NP1, dims = 1:9)

DimPlot(NP1, reduction = "umap", pt.size = 0.5,  label = TRUE) + NoLegend()




# Annotation 

new.cluster.ids <- c("Basal1", "ML1", "LP1", "ML2", "ML3", "ML4", "LP2", 
    "ML5", "Basal2","Rare1", "Rare2", "Rare3")
names(new.cluster.ids) <- levels(NP1)
NP1 <- RenameIdents(NP1, new.cluster.ids)
DimPlot(NP1, reduction = "umap", label = TRUE, pt.size = 0.5) + NoLegend()


```
# Save the intermediate clusters 
```{r}
saveRDS(NP1, file = "~/data/NP1_tutorial.4.rds")
```
